### 2021-02-01 FS-CRB-HR-report1.tex
added rept 1

\documentclass[12pt,letterpaper,english,bibliography=totocnumbered,abstract=on]{scrartcl}

\usepackage{indentfirst}
\usepackage{appendix}
\usepackage{fullpage}
%\usepackage{subfiles}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{color}
\usepackage{babel}
\usepackage{verbatim}
\usepackage[unicode=true,pdfusetitle,
bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
breaklinks=true,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]
{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,urlcolor=blue}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{threeparttable}
\usepackage[table]{xcolor}
\usepackage{csquotes}
\usepackage{soul} % for hiliting text: \hl
% old style is authoryear
\usepackage[backend=biber, style=numeric, maxbibnames=99]{biblatex}
\addbibresource{mylibrary.bib}
\addbibresource{CRB.bib}

\usepackage[disable]{todonotes}

% Prevent page breaks within paragraphs
% https://tex.stackexchange.com/questions/21983/how-to-avoid-page-breaks-inside-paragraphs
\widowpenalties 1 10000

\begin{document}
\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 1}
\title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
\author{Aubrey Moore, University of Guam}
%\date{Submitted December 30, 2020\\Revised January 28, 2021}
\maketitle
\begin{description}	
	\item[GRANTEE:] Aubrey Moore, University of Guam 
	\item[GRANT YEAR:] 2020
	\item[GRANT NUMBER:] 20-DG-11052021-227
	\item[GRANT PROGRAM:] Forest Health Protection
	\item[GRANT EXPIRATION DATE:] 2021-05-30
	\item[DATES COVERED BY THIS REPORT:] 2020-06-17 through 2020-12-31
	\item[GRANT STATUS:] Active
\end{description}

\begin{footnotesize}
	\todo{change url}
\url{https://github.com/aubreymoore/FY19-PPA-Report-1/raw/master/PPA19-report2.pdf}
\end{footnotesize}

\newpage{}
\tableofcontents{}

\newpage
\listoftodos

\newpage

\section{OBJECTIVES AND SPECIFIC ACTIVITIES}

%List the objectives individually that were included in the grant narrative.  Under those objectives include specific activities that occurred during the report timeframes.

The objective of this grant project is to evaluate harmonic radar as an alternative to radio tracking for CRB breeding site detection. Detection of CRB breeding sites is essential for CRB control and eradication.

In a previous Forest Service feasability study, we successfully tracked CRB to cryptic breeding sites using miniature radio transmitters. However, the high cost of radio tags (>\$100 each) and short shelf life and field life of nonreplacable batteries (about 1 month and one week, respectively) preclude radio tracking for routine CRB breeding site detection make this technique too expensive for routine surveys.

A promising alternative is harmonic radar technology. Harmonic radar tags do not require a battery. They are inexpensive (about \$1 each) and they have unlimited shelf-life and field life. We are evaluating hand-held harmonic radar equipment and tags manufactured by RECCO in Sweden. Rescuers use the hand-held units to rapidly locate victims wearing tags sown into their clothing. RECCO technology has been used to track other insects but has not yet been tried with CRB.

Our objective is not to track CRB tagged beetles in real-time, but to discover the end points of tags several days or even weeks after release of the tagged beetles.We anticipate that tags will accumulate at breeding sites.

Here is the plan from the \textit{verbatim} from the approved grant proposal:

\medskip
\colorbox[gray]{0.9}{\parbox{\textwidth}{
	\textbf{Schedule of Activities (2020)}
	\begin{description}
		\item[March-April:] Tag fabrication and testing at EMU.
		\item[April-early May:] Capture, flight test, and mark CRB at UoG.
		\item[May:] Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period).
		\item[June-August:] Data analysis.
		\item[August-December:] Manuscript(s) and final report preparation.  Discuss findings with state agencies.  Make presentations at scientific meetings.  Plan further research with cooperators to implement findings in monitoring and control efforts.
	\end{description}
}}
\medskip

Progress to date includes:
\begin{itemize}
\item Procurement of harmonic radar equipment and tags
\item Tag fabrication (soldering antennae to diodes) in Matt Siderhurst's lab.
\item Preliminary field testing of equipment on Guam.
\end{itemize}

The most important activity is \textbf{Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period)}. The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar (similar to what we did in the radio-tracking feasibility study). This activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}. Otherwise we are prepared to proceed.

\section{OUTPUTS}

%Each Output listed in your grant narrative addressed individually.
%
%1) Output 1: list the output and what was accomplished.
%2) Output 2: list the output and what was accomplished.
%3) Etc.
%
%If output includes positions funded: (who, how long, type of work)
%
%If output includes Acres treated, include target species, number of acres and location 
%
%If output includes Acres surveyed, include target species, number of acres and location 
%
%Description and dollar value of equipment purchased, 
%
%Number of personnel trained.

Nothing to report.

\section{MONITORING \& EVALUATION}

%If post-treatment monitoring has been completed, provide the results, especially results that show effectiveness of treatments.  
%
%List any project evaluations that took place to determine whether goals and Statewide Strategies are being met.

Nothing to report.

\section{BUDGET EXPENDITURES}

%Include budget table showing the expenditures (Personnel/salary costs, supplies purchased, contracts, etc.) that occurred during the reporting period

\begin{tabular}{lrrl}
	\hline
	Category & Budget & Spent & Note \\
	\hline 
	Equipment and supplies & \$8,000 & \$3,462 & Harmonic radar and tags \\ 
	Travel & \$12,000 & \$0 \\ 
	Admin. fee & \$3,000 & \$0 \\ 
	\hline 
\end{tabular}

\label{impediments}
\section{PROBLEMS ENCOUNTERED THIS REPORTING PERIOD}

%Explain delays, adverse conditions or changed costs that significantly impair the ability to meet grant objectives.  If necessary, prepare a separate formal request for an extension of the grant period:

Progress on this project was impeded by COVID-19 travel restrictions which prevented collaborators from visiting Guam to participate in field work. Further delayed by delay was caused by Government of Guam \textit{stay at home} orders. The University of Guam was officially closed from March 20 to May 10 2020 and again from August 16 2020 to January 15 2021.

\section{CHANGES PLANNED}

% (If the scope of the objectives would change, or if more that 10\% of the total grant budget would change object class categories, prepare a separate formal request to amend the grant):

Nothing to report.

\section{CIVIL RIGHTS}

% - any activity that demonstrates Title VI compliance with civil rights requirements, such as documentation of outreach to underserved groups, public education documents in languages other than English, etc

Nothing to report.

\section{ATTACHMENTS}

% (photos of activities are very helpful, also electronic copies of survey and treatment maps, copies of developed brochures or posters, etc.):

None.

\section{PLANS}

% for work to be performed during next reporting period:

Nothing to report.

\end{document}

### 2021-02-01 FS-CRB-HR-report1.tex
final

@@
 \begin{footnotesize}
-	\todo{change url}
-\url{https://github.com/aubreymoore/FY19-PPA-Report-1/raw/master/PPA19-report2.pdf}
+\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report1.pdf}
 \end{footnotesize}
@@
 
-In a previous Forest Service feasability study, we successfully tracked CRB to cryptic breeding sites using miniature radio transmitters.
-However, the high cost of radio tags (>\$100 each) and short shelf life and field life of nonreplacable batteries (about 1 month and one week, respectively) preclude radio tracking for routine CRB breeding site detection make this technique too expensive for routine surveys.
+Previously, we successfully tracked CRB to cryptic breeding sites using miniature radio transmitters.
+However, the high cost of radio tags (>\$100 each) and short shelf life and field life of nonreplacable batteries (about 1 month and one week, respectively) make this technique too expensive for routine surveys to detect CRB breeding sites.
 
-A promising alternative is harmonic radar technology.
+A promising alternative technology is harmonic radar.
 Harmonic radar tags do not require a battery.
@@
 
-Our objective is not to track CRB tagged beetles in real-time, but to discover the end points of tags several days or even weeks after release of the tagged beetles.We anticipate that tags will accumulate at breeding sites.
+Our objective is not to track CRB tagged beetles in real-time, but to discover the end points of tags several days or even weeks after release of tagged beetles.
+We anticipate that tags will accumulate at breeding sites.
 
-Here is the plan from the \textit{verbatim} from the approved grant proposal:
+Here is the plan from the approved grant proposal:
 
@@
 The most important activity is \textbf{Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period)}.
-The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar (similar to what we did in the radio-tracking feasibility study).
-This activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}.
+The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar.
+(We are following an approach similar to what we did in the radio-tracking feasibility study).
+This fieldwork activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}.
 Otherwise we are prepared to proceed.
@@
 Further delayed by delay was caused by Government of Guam \textit{stay at home} orders.
+
 The University of Guam was officially closed from March 20 to May 10 2020 and again from August 16 2020 to January 15 2021.

### 2021-02-22 experiments/techreport/techreport.tex
added expt and tch rept.

\documentclass[12pt,letterpaper,english,bibliography=totocnumbered, abstract=on]{scrartcl}

\usepackage{indentfirst}
\usepackage[titletoc]{appendix}
%\usepackage{fullpage}
%\usepackage{subfiles}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{color}
\usepackage{babel}
\usepackage{verbatim}
\usepackage[unicode=true,pdfusetitle,
bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
breaklinks=true,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]
{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,urlcolor=blue}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{threeparttable}
\usepackage[table]{xcolor}
\usepackage{csquotes}
\usepackage{soul} % for hiliting text: \hl

\usepackage[backend=biber, sorting=none]{biblatex}
%\setlength\bibitemsep{2\itemsep}
%\addbibresource{mylibrary.bib}
\addbibresource{CRB.bib}

\usepackage{pdfpages}
\usepackage{float} % Allows use of H to place floats

\usepackage{pgfgantt}

\usepackage{framed}
\usepackage{todonotes}

% Prevent page breaks within paragraphs
% https://tex.stackexchange.com/questions/21983/how-to-avoid-page-breaks-inside-paragraphs
\widowpenalties 1 10000

\begin{document}

\titlehead{Technical Report}

\title{Using a Cell Phone to as a Datalogger for Handheld Harmonic Radar}

\author{Aubrey Moore PhD\\University of Guam College of Natural and Applied Sciences}

\date{\today}

\maketitle

%\begin{footnotesize}
%	\url{https://xxx}
%\end{footnotesize}

\clearpage

For scientific applications, such as tracking tags attached to insects, it is desirable to log output from the hand-held RECCO harmonic radar device along with timestamp, GPS coordinates, direction of signal, intensity of signal, etc. Here, I report on my attempts use a cell phone as a datalogger.

When the RECCO detects a tag, it alerts the user with an audible beeps. The intensity of these beeps indicates the strength of the returned signal and the direction of the tag is in line with the Yagi antenna.

\section{Materials and Methods}

\subsection{Setting up a Smart Phone}

\subsubsection{Hardware}

I used a Samsung Galaxy 5 smart phone. Beeps were recorded by connecting the phone's audio jack to the headphone plug RECCO. It was necessary to build a simple interface circuit before the phone would record from the audio jack (Fig. \ref{fig:adapter}). This circuit was integrated with a custom made plug for the RECCO headphone jack (Figs. \ref{fig:recco1},\ref{fig:recco2}).

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{images/adapter}
	\caption{Line-in adaptor circuit for cell phones. Source: \url{https://warmplace.ru/docs/mobile_audio_input/}}
	\label{fig:adapter}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth,angle=-90]{images/recco1}
	\caption{RECCO hand-held harmonic radar with smart phone attached.}
	\label{fig:recco1}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth,angle=-90]{images/recco2}
	\caption{RECCO hand-held harmonic radar with smart phone attached.}
	\label{fig:recco2}
\end{figure}

The phone was attached on top of the RECCO's handle using a ball mount and phone holder (Figs. \ref{fig:recco1},\ref{fig:recco2}).

\subsubsection{Software}

I used a free \href{https://www.arduino.cc/education/science-journal}{Arduino Science Journal app}. This app allows selection of various sensors for recording experimental data.  In this case, I used the sound intensity sensor, and the compass direction sensor.

\subsection{Field Test}

Prior to connecting the phone to the REVCCO, I \href{https://www.howtogeek.com/519142/how-to-calibrate-the-compass-on-android-to-improve-device-location-accuracy/}{calibrated the compass sensor}.

A RECCO test tag was clamped to a tripod at about 1.5 m above ground at one end of a small field Fig. (\ref{fig:target}). At the other end of the field, 70 m away (Fig.\ref{fig:screenshot-from-2021-02-22-08-47-24}), I made several counter clockwise rotations while holding the RECCO with smart phone as in Fig. \ref{fig:recco2}.

\begin{figure}[p]
	\centering
	\includegraphics[width=0.4\textwidth, angle=-90]{"images/target"}
	\caption{RICCO test target containing a harmonic radar tag.}
	\label{fig:target}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{"images/Screenshot from 2021-02-22 08-47-24"}
	\caption{Test setup. The line shows the 70m distance between the RECCO harmonic rader device on the left and the test target tag on the right.}
	\label{fig:screenshot-from-2021-02-22-08-47-24}
\end{figure}

\section{Analysis and Results}

The CSV data file recorded by the Arduino Science Journal app was downloaded to my laptop computer and analysed using a Jupyter notebook (see Figs. \ref{fig:compass},\ref{fig:decibels},\ref{fig:polar0},\ref{fig:polar75}).

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{images/compass}
	\caption{Compass direction.}
	\label{fig:compass}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{images/decibels}
	\caption{Sound intensity.}
	\label{fig:decibels}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{images/polar0}
	\caption{Polar plot of sound intensity versus compass direction.}
	\label{fig:polar0}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=0.7\linewidth]{images/polar75}
	\caption{Polar plot of sound intensity versus compass direction.}
	\label{fig:polar75}
\end{figure}

\section{Discussion}

All of the signals above 75 dB were from the east quadrant, centered on the location of the test tag (Fig. \ref{fig:polar75}).

The beeps are on top of a very loud (60 dB to 70 dB)white noise background. It should be possible to increase the  signal-to-noise ratio significantly by recording the raw waveform (as a WAV file) instead of calculating decibels integrated over all frequencies. A spectral filter could then be used to ignore everything except the beep frequency.

\end{document}

### 2021-09-30 FS-CRB-HR-report2.tex
added rept 2

--- 2021-02-01 FS-CRB-HR-report1.tex
@@
 \begin{document}
-\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 1}
+\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 2}
 \title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
@@
 	\item[GRANT PROGRAM:] Forest Health Protection
-	\item[GRANT EXPIRATION DATE:] 2021-05-30
-	\item[DATES COVERED BY THIS REPORT:] 2020-06-17 through 2020-12-31
+	\item[GRANT EXPIRATION DATE:] 2022-12-31 (extended)
+	\item[DATES COVERED BY THIS REPORT:] 2021-01-01 through 2021-06-01
 	\item[GRANT STATUS:] Active
@@
 \begin{footnotesize}
-\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report1.pdf}
+\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report2.pdf}
 \end{footnotesize}
@@
 \newpage
+
+% TODO: \usepackage{graphicx} required
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{drone}
+	\caption{RECCO harmonic radar device suspended from an agricultural drone.}
+	\label{fig:drone}
+\end{figure}
+
 \listoftodos
@@
 
-The most important activity is \textbf{Conduct field releases and tracking of tagged CRB (two week intensive fieldwork period)}.
+The most important activity is to conduct field releases and tracking of tagged CRB over a two week intensive fieldwork period.
 The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar.
 (We are following an approach similar to what we did in the radio-tracking feasibility study).
-This fieldwork activity has not yet been scheduled because of COVID-19 travel restrictions \ref{impediments}.
+This fieldwork activity has not yet been scheduled because of COVID-19 travel restrictions (See section \ref{impediments} for details.).
 Otherwise we are prepared to proceed.
+
+\subsection{Methods Development}
+
+\textbf{The RECCO hand-held harmonic radar device we are using is designed for ground-based location of tags.
+We are collaborating with Dr.
+Glenn Dulla, Guam Department of Agriculture, to test the idea of mounting the device on a drone to allow efficient downward-looking scans over large areas.
+The idea is to mount the RECCO device on a drone which is programmed to fly a defined search path.
+A digital recorder cabled to the RECCO's headphone jack records the audio signal from the device and the drone's GPS  records location.
+Georeferencing is done by matching timestamps in the audio recording with timestamps in the GPS data.
+Postprocessing generates a map of signal strength with peaks indicating probable tag locations requiring confirmation by ground searches.
+}
+
+\textbf{Preliminary tests are promising. To date we have:
+\begin{itemize}
+	\item Used digital recordings of the RECCO's audio output to locate tags (Technical report)
+	\item Flown the RECCO harmonic radar to a drone (Fig. \ref{fig:drone})
+\end{itemize}
+}
 
@@
 
+\section{PROBLEMS ENCOUNTERED THIS REPORTING PERIOD}
 \label{impediments}
-\section{PROBLEMS ENCOUNTERED THIS REPORTING PERIOD}
 
@@
 The University of Guam was officially closed from March 20 to May 10 2020 and again from August 16 2020 to January 15 2021.
+
+\textbf{COVID travel restrictions during the current reporting period did not allow Dr.
+Siderhurst and students to visit Guam to perform the planned field work.
+Guam is currently experiencing high COVID incidence predominantly caused by the D variant.
+We are currently targeting spring break 2022 as the time period for field work.}
 

### 2021-09-30 FS-CRB-HR-report2.tex
added rept 2

@@
 
+\usepackage{graphicx}
+
 % Prevent page breaks within paragraphs
@@
 
+\textbf{Note to Reader:} Significant changes from the previous report are indicated by \textbf{bold-faced text}.
+
+\bigskip
+
 \begin{footnotesize}
@@
 
-% TODO: \usepackage{graphicx} required
 \begin{figure}
@@
 	\includegraphics[width=0.7\linewidth]{drone}
-	\caption{RECCO harmonic radar device suspended from an agricultural drone.}
+	\caption{RECCO harmonic radar device suspended from an agricultural drone piloted by Dr. Glenn Dulla, Guam Department of Agriculture.}
 	\label{fig:drone}
@@
 
-\listoftodos
-
-\newpage
+\clearpage
 
@@
 \begin{itemize}
-	\item Used digital recordings of the RECCO's audio output to locate tags (Technical report)
-	\item Flown the RECCO harmonic radar to a drone (Fig. \ref{fig:drone})
+	\item Used digital recordings of the RECCO's audio output to locate tags (See Technical Report: Using a Cell Phone as a Datalogger for Handheld Harmonic Radar \url{https://github.com/aubreymoore/Harmonic-Radar/blob/master/experiments/techreport/techreport.pdf})
+	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
 \end{itemize}

### 2021-09-30 FS-CRB-HR-report2.tex
added rept 2

@@
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
-breaklinks=true,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]
+pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]
 {hyperref}

### 2022-02-16 FS-CRB-HR-report3.tex
report3

--- 2021-09-30 FS-CRB-HR-report2.tex
@@
 \begin{document}
-\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 2}
+\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 3}
 \title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
@@
 	\item[GRANT EXPIRATION DATE:] 2022-12-31 (extended)
-	\item[DATES COVERED BY THIS REPORT:] 2021-01-01 through 2021-06-01
+	\item[DATES COVERED BY THIS REPORT:] 2021-06-01 through 2021-12-31
 	\item[GRANT STATUS:] Active
@@
 \begin{footnotesize}
-\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report2.pdf}
+\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report3.pdf}
 \end{footnotesize}
@@
 
-\textbf{The RECCO hand-held harmonic radar device we are using is designed for ground-based location of tags.
+The RECCO hand-held harmonic radar device we are using is designed for ground-based location of tags.
 We are collaborating with Dr.
@@
 Postprocessing generates a map of signal strength with peaks indicating probable tag locations requiring confirmation by ground searches.
-}
 
-\textbf{Preliminary tests are promising. To date we have:
+Preliminary tests are promising:
 \begin{itemize}
@@
 	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
+	\item \textbf{Based on initial flight tests we determined that the audio data recorder we were using was too heavy for use on the drone.
+We purchased a much smaller and light audio data recorder.
+Walk tests with this recorder show that we can locate harmonic radar tags (Fig.
+\ref{fig:211109003}).}
 \end{itemize}
-}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\linewidth]{211109_003.png}
+	\caption{Audio output from a RECCO harmonic radar device recorded using a Zoom F2 field recorder.
+This was a walk test along a linear transect in which a harmonic radar tag had been placed.
+Location of the tag in the recording is clear.}
+	\label{fig:211109003}
+\end{figure}
+
+\clearpage
 
@@
 
-Nothing to report.
-
+\textbf{On December 18, Aubrey Moore made a presentation on the technology used in this project at a meeting of the Guam Beekeepers Association (GBA).
+Members of the GBA and the Guam Department of Agriculture are interested in using harmonic radar to track greater banded hornets (GBH), \textit{Vespa tropica}, to their nests so that they can be destroyed.
+GBH, a recently arrived invasive species which raids honeybee hives, is a serious threat to Guam's nascent beekeeping industry.
+}
 \section{MONITORING \& EVALUATION}
@@
 	\hline 
-	Equipment and supplies & \$8,000 & \$3,462 & Harmonic radar and tags \\ 
+	Equipment and supplies & \$8,000 & \$316.93 & Miniturized audio field recorder and hardware \\ 
 	Travel & \$12,000 & \$0 \\ 
@@
 
-\textbf{COVID travel restrictions during the current reporting period did not allow Dr.
+COVID travel restrictions during the current reporting period did not allow Dr.
 Siderhurst and students to visit Guam to perform the planned field work.
 Guam is currently experiencing high COVID incidence predominantly caused by the D variant.
-We are currently targeting spring break 2022 as the time period for field work.}
+We are currently targeting spring break 2022 as the time period for field work.
+
+\textbf{As of mid February 2022, Guam is still has extremely high local transmission of COVID due to delayed arrival of the OMICRON variant.
+However, quarantine for fully vaccinated visitors from the mainland is no longer required.
+A decision on when Dr.
+Siderhurst and students experienced with using harmonic radar to track insects will visit Guam to complete planned field work is still pending.}
 

### 2022-10-29 FS-CRB-HR-report4.tex
catchup

--- 2022-02-16 FS-CRB-HR-report3.tex
@@
 \begin{document}
-\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 3}
+\titlehead{US Forest Service Forest Health Protection Grant Progess Report 4}
 \title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
@@
 	\item[GRANT EXPIRATION DATE:] 2022-12-31 (extended)
-	\item[DATES COVERED BY THIS REPORT:] 2021-06-01 through 2021-12-31
+	\item[DATES COVERED BY THIS REPORT:] 2022-01-01 through 2022-06-30
 	\item[GRANT STATUS:] Active
@@
 \begin{footnotesize}
-\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report3.pdf}
+\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report4.pdf}
 \end{footnotesize}
@@
 \item Preliminary field testing of equipment on Guam.
+\item \textbf{Published our research idea as a preprint in the Research Ideas and Outcomes Journal (\href{https://doi.org/10.3897/rio.8.e86422}{Moore and Siderhurst 2022}).
+This article will be peer reviewed.}
 \end{itemize}
 
-The most important activity is to conduct field releases and tracking of tagged CRB over a two week intensive fieldwork period.
-The intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar.
-(We are following an approach similar to what we did in the radio-tracking feasibility study).
+The most important activity in our work plan is to conduct field releases and tracking of tagged CRB over a two week intensive fieldwork period.
+This intensive fieldwork will be performed on Guam in collaboration with Matt Siderhurst's team of students who have experience in tracking insects with harmonic radar.
+(We are following an approach similar to what we did in the radio-tracking feasibility study \href{https://academic.oup.com/ee/article/46/1/92/2633399?login=false}{Moore et al.
+2017}).
 This fieldwork activity has not yet been scheduled because of COVID-19 travel restrictions (See section \ref{impediments} for details.).
 Otherwise we are prepared to proceed.
+\textbf{Our field trial is now planned for July 2022.}
 
@@
 	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
-	\item \textbf{Based on initial flight tests we determined that the audio data recorder we were using was too heavy for use on the drone.
+	\item Based on initial flight tests we determined that the audio data recorder we were using was too heavy for use on the drone.
 We purchased a much smaller and light audio data recorder.
-Walk tests with this recorder show that we can locate harmonic radar tags (Fig.
-\ref{fig:211109003}).}
+Walk tests with this recorder show that we can locate harmonic radar tags by correlating audio recordings with GPS data (Fig.
+\ref{fig:211109003}).
+	\item \textbf{Recent drone flight.}
 \end{itemize}
@@
 
-COVID travel restrictions during the current reporting period did not allow Dr.
-Siderhurst and students to visit Guam to perform the planned field work.
-Guam is currently experiencing high COVID incidence predominantly caused by the D variant.
-We are currently targeting spring break 2022 as the time period for field work.
-
-\textbf{As of mid February 2022, Guam is still has extremely high local transmission of COVID due to delayed arrival of the OMICRON variant.
+\textbf{COVID travel restrictions during the current reporting period did not allow Dr.
+Siderhurst and students to visit Guam to perform the planned field work on schedule.
+Guam still has extremely high local transmission of COVID due to delayed arrival of the OMICRON variant.
 However, quarantine for fully vaccinated visitors from the mainland is no longer required.
-A decision on when Dr.
-Siderhurst and students experienced with using harmonic radar to track insects will visit Guam to complete planned field work is still pending.}
+Travel arrangements have been made for Dr.
+Siderhurst and students experienced with using harmonic radar to track insects to visit Guam during July 2022 to complete planned field trial.}
 

### 2022-10-29 FS-CRB-HR-report4.tex
safety

@@
 This article will be peer reviewed.}
+\item \textbf{Continued to investigate the idea of mounting our harmonic radar transceiver on an aerial drone to facilitate automated surveys over large areas (See Methods Development section below).
+}
 \end{itemize}
@@
 \begin{itemize}
-	\item Used digital recordings of the RECCO's audio output to locate tags (See Technical Report: Using a Cell Phone as a Datalogger for Handheld Harmonic Radar \url{https://github.com/aubreymoore/Harmonic-Radar/blob/master/experiments/techreport/techreport.pdf})
+	\item Used digital recordings of the RECCO's audio output to locate tags (See Technical Report: \href{https://github.com/aubreymoore/Harmonic-Radar/raw/master/experiments/techreport/techreport.pdf}{Using a Cell Phone as a Datalogger for Handheld Harmonic Radar})
 	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
@@
 \ref{fig:211109003}).
-	\item \textbf{Recent drone flight.}
+	\item \textbf{On June 9, 2022 we flew a drone with our harmonic radar transceiver suspended beneath it.
+Audio output from the transceiver was recorded on a small audio data recorder.
+By correlating audio data with the drone's GPS log, we were able to locate harmonic detector tags on the ground (See \href{https://github.com/aubreymoore/Harmonic-Radar/raw/master/experiments/2022-06-09-drone/techreport.pdf}{technical report}).}
 \end{itemize}
@@
 
-\textbf{On December 18, Aubrey Moore made a presentation on the technology used in this project at a meeting of the Guam Beekeepers Association (GBA).
+On December 18, Aubrey Moore made a presentation on the technology used in this project at a meeting of the Guam Beekeepers Association (GBA).
 Members of the GBA and the Guam Department of Agriculture are interested in using harmonic radar to track greater banded hornets (GBH), \textit{Vespa tropica}, to their nests so that they can be destroyed.
 GBH, a recently arrived invasive species which raids honeybee hives, is a serious threat to Guam's nascent beekeeping industry.
-}
+
 \section{MONITORING \& EVALUATION}
@@
 
-\section{PROBLEMS ENCOUNTERED THIS REPORTING PERIOD}
+\section{PROBLEMS ENCOUNTERED DURING THIS REPORTING PERIOD}
 \label{impediments}

### 2023-03-11 harmonic-radar_article/response-to-review-comments.md
catchup

# Harmonic Detection RIO Article Peer Review

# Reviewer 1

## Comments

I’d suggest title should be Proposal for detecting…..  as it is a proposal, so being just in the proposal stage, there is no data to give.

The testing done to indicate the CRB biotype on Guam is resistant to OrNV may well have been done using OrNV produced in Heteronychus arator tissue culture, a different species. But the possibility has been suggested that the OrNV produced in this way may have mutated and evolved so as to adapt to infecting Heteronychus cells, given OrNV’s very short generation time in this culture, but at the same time its ability to attack cells of the CRB Oryctes rhinoceros may become reduced or even cancelled as it is no longer in contact with CRB cells, the intended target. Nobody knows at this stage. Reference is:  Bedford, G.O. (2018) Possibility of Evolution in Culture of the Oryctes Nudivirus of the Coconut Rhinoceros. Beetle Oryctes rhinoceros (Coleoptera: Scarabaeidae: Dynastinae). Advances in Entomology, 6, 27-33. https://doi.org/10.4236/ae.2018.61004.           Also, recent publications have shown CRB biotypes on Solomon Islands, & Palau, similar to the Guam biotype, have been found to be infected with or carrying OrNV. I can give references if needed.

Could include a bit more mention (as mentioned above in response to question re citing literature pertinent) of possibility of OrNV mutating, evolving & adapting in Heteronychus tissue culture & this affecting its ability to infect CRB Oryctes rhinoceros biotypes.

## Response

The title has been changed to "Proposal for detecting…."  as suggested by the reviewer.

The reviewer points out that it is not known why *Oryctes rhinoceros* biotype G is apparently resistant to *Oryctes* nudivirus and suggests one of several possible reasons being investigated. However, a discussion of the reasons for this biocontrol failure are not relevant to this article.

<hr>

# David Hall

## Comments

This is an interesting concept and generally well-written paper.

However, it is completely lacking in any quantitative data as to how many insects could be reasonably tagged (cost, time, etc) and then what percentage of these beetles could be effectively tracked. The RECCO transponder has relatively short range (not specified here) and so it seems to me that once the beetles have been released it would be necessary to carry out very extensive surveys to find them again. These surveys would thus be restricted to what field experts judge to be likely breeding sites, and so it is not clear to me what the added value would be of actually locating a tagged beetle at one of these sites.

There is also no real information on what would then be done to these sites and how many it would require to make an efective dent in the beetle population and damage.

## Response

Quantitative data are lacking because this is a Research Idea article. We have just begun field trials to test the feasability of our idea and will report data results in a subsequent Research Outcomes article.

However, we were able to update the draft to include cost and time estimates for harmonic tags and detection range estimates as requested by the reviewer.

AUTOMATED SURVEYS
"The RECCO transponder has relatively short range (not specified here) and so it seems to me that once the beetles have been released it would be necessary to carry out very extensive surveys to find them again."
In the revised text, we suggest "Searches may be highly automated by mounting a HR transceiver equipped with a data logger on an aerial drone. The drone will fly programmed search paths close to the ground with the HR beam pointing downwards, thus compensating for the relative short detection range of the HR transceiver."

SANITATION METHODS
"There is also no real information on what would then be done to these sites and how many it would require to make an efective dent in the beetle population and damage." A sentence listing sanitation methods and a reference providing details have been added.

<hr>

# Widihastuty

## Comments

My comment

1. please explained the tools that used to detect the breeding site of O. rhinoceros, how work the tools

2. data not sufficient to support the tools

3. conclusions does not explained the results of the detection that has been succesfully carried out by the tool

4. reference too less

## Response

1. The text describes 4 methods for detecting CRB breeding sites. Our proposed use of harmonic radar for finding CRB breeding sites is described in detail.
2. This is a Research Idea article. We have just begun field trials to test the feasability of our idea and will report results in a subsequent Research Outcomes article.
3. Same as above item.
4. Additional references have been cited.
